Universal regime transition for water resources utilization in developing river basins

### 2020-09-26 tex/main/manuscript.tex
update abstract

% 文章标题：“流域尺度的水资源利用体系：过渡框架和发展困局”
\title{Water resource utilization regimes at a basin scale: transition framework and development traps}

@@

%tag 摘要
\begin{abstract}
	From agrarian to industrial societies, humans have harnessed ecosystems over rivers basins around the world, while using water resources as blood that sustains social developments. 
	However, there is still lacking of effective method to detect regime shifts of water utilization linking to social changes, with much fewer attempts to develop theoretical models to explain their transition phases as well. 
	Here, by integrating three different dimensions of water utilization (stress, tendentiousness and pattern), we develop an Integrated Water Resources Utilization (IWRU) Index at a basin scale to detect regime shifts. 
	By applying this index to the Yellow River Basin, China, we summarized a general transition framework of regimes, which gives a sketch of relationships between human societies and their water utilization. 
	Based on the framework, transition phases within the three regimes and accompanying development traps can be well understood, as a useful guideline for the Yellow River to develop in a coordinated way.
\end{abstract}

@@

% Tag 引言第三段
% 前人已经从不同维度刻画了人水关系.
Development of societies by using water resources has been going on for at least thousands of years. Although its regime shifts and transition phases are not fully understood yet, water utilization has been depicted and studied from different dimensions.
% 首先，因为水资源的稀缺性和全球用水量的增加，受到最广泛关注的是人类社会面临的水资源压力。
Firstly, since water resources are scarce, the most widespread concern is the rising stresses on human societies to use water resources.
Even though the stocks of water in artificial reservoirs are helpful to water resources availability, greater water utilization stresses had become a major constraint to development, because of significant increment in water withdrawals and larger shares of inflexible water utilization during the last century.
\cite{postelHumanAppropriationRenewable1996, greveGlobalAssessmentWater2018a, qinFlexibilityIntensityGlobal2019}
% 其次，随着工业迅猛发展和生态建设的需要，社会对水资源的利用的倾向性也发生了转变。
Secondly, as the need of industrial and ecological developments, tendentiousness of water utilization changed with. 
Despite a major water utilization of agricultural irrigation dominating most river basins, there are noticeable growths and preferential tendentiousness in the economy profits and water consumption regarding industry or services, leading potential conflicts between different sectors.
\cite{liuWaterScarcityAssessments2017, florkeWaterCompetitionCities2018}
% 最后，由于水的可用性本质上是区域问题，水资源利用的格局也很重要
Thirdly, since water distribution and utilization are inherently basinal concerns, patterns of also play an important role.
% 全球总的取水量很少，但缺水地区很多
Although only 10\% of available water is withdrawn on global average, about 30\% of population live in highly water-stressed areas,
where dominated sectors of water utilization are various as well. 
\cite{wadaWedgeApproachWater2014, okiGlobalHydrologicalCycles2006}
% 此外，人类活动还在改变这一格局
In addition, human activities are still changing this pattern, since positive impacts caused by human interventions mostly occur in upper regions whereas aggravated water resources downstream, in many basins around the world.
\cite{veldkampWaterScarcityHotspots2017}
% 将三种视角结合，就是“水资源利用体系”。
Although existing researches have evaluated the aspects of water resource utilization from these different dimensions, we still cannot obtain a coherent understanding of regime shifts regard to social development and water utilization, without integrating them.

@@

\section*{Results}
\subsection*{Detection of Water utilization regimes}
% 这一节主要展示IWRU的变化趋势和WUR的划分
With two significant points detected, the trend of IWRU index are split into three periods, whose slopes are various and mainly contributed by different dimensions (stress, tendentiousness or pattern of water utilization, see Methods) (Figure~\ref{fig:IWRU}).
\begin{figure}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/index.pdf}
	\caption{Changes of the IWRU index. 
	\textbf{A,} with two change points in 1978 and 1994, three periods were detected in trend of the IWRU.
	\textbf{B,} changes of IWRU in three periods have various slopes, while the second period have a negative growths rate.
	\textbf{C,} changes of the IWRU within three certain periods, which have different main contributors.
	}
	\label{fig:IWRU}
\end{figure}
% 接下来分句介绍每个阶段的特征
% 第一阶段
In the first period (P1, 1965-1978), the IWRU index had a rapidly increasing and the lightening of water stresses made the most striking contribution (+0.722), while tendentiousness and pattern of the water utilization had slight negative contribution (-0.048 and -0.09 respectively).
% 第二阶段
In the second period (P2, 1979-1994), the IWRU index experienced a slight drop, despite positive contributions of tendentiousness and pattern of water utilization (+0.352 and +0.279 respectively), because of increasing stresses on water resource playing a larger negative role (-0.636). 
% 第三阶段
However, as the further increasing of positive contributions of water utilization tendentiousness (+0.485) and pattern (+0.515), and decelerations of water stresses (-0.344, 46\% less than P2) in the third period (P3, 1995-2013), a positive growth of the IWRU returned.
% 总而言之，每个阶段都由水资源利用的不同维度提供最大的正向作用
As a result, each period has a different most striking positive contributor to IWRU: P1 is stress; P2 is tendentiousness; and P3 is pattern.

% 用水三个维度的组合呈现出阶段特征明显
Combining these three dimensions' net contribution to IWRU further, ratios of the contributions of the three dimensions clustered clearly by different time periods, indicating three regimes (Figure~\ref{fig:phases}).
\begin{figure}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/phases.pdf}
	\caption{Combination of three dimensions (S: stresses; T: tendentiousness; P: pattern) in different periods. 
	Size of the points denoting average of absolute value of each indicator: the mean of the P1 phase is 0.10, while 0.14, 0.19 in P2 and P3.
	The red indicator line in this ternary plot denotes 1:1 contributions between tendentiousness (T) and patterns (P).
	Two key change points (1978, 1994), along with the beginning (1965) and the ending (2013) of research period, are labelled.}
	\label{fig:phases}
\end{figure}
% 第一阶段到第二阶段
At the very beginning (1965) and throughout the whole P1, water utilization regime dominated by high stress. After then, it experienced a shift to low stress since 1978, with a change in the proportion of contributions between tendentiousness and pattern, too.
% 第三阶段集中
Finally, the three dimensions' contribution were much similar in P3, making the points highly concentrated at the centre of the ternary diagram for that period.

%tag 结果2
\subsection*{Differences between water utilization regimes}
% 在不同的维度下各阶段进行对比，可以发现不同的水资源利用体系间存在明显差异。
The differences between the water utilization regimes are reflected in changes of all three dimensions.
\begin{figure*}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/dimensions.pdf}
	\caption{
		Changes in different dimensions of water resources utilization regimes and their main contributors.
		\textbf{A,} changes of water utilization stress, indicated by unstandardized scarcity-flexibility-variability water stresses index (SFV-index).
		\textbf{B,} changes of water utilization tendentiousness, indicated by non-provisioning water shares.
		\textbf{C,} changes of water utilization pattern, indicated by unstandardized distribution information entropy index (see supplementary: Methods. S3 for details).
		\textbf{D,} Main impact factors to water utilization stresses in each period or region, and their contributions to changes of unstandardized SFV-index.
		\textbf{E,} Main impact water uses to water utilization tendentiousnesses, and their contributions to changes of non-provisioning water shares.
		\textbf{F,} Main impact factors to water utilization patterns, and their contributions to changes of related unstandardized distribution information entropy index (see supplementary: Methods. S3 for details).
		}
	\label{fig:dimensions}
\end{figure*}
% 从阶段一到阶段二，水资源压力的变化最为明显
Moving from the regime in P1 to P2, the most striking change is the reversal of the trend in water utilization stress, which is determined by a combination of scarcity, flexibility and variability (Figure~\ref{fig:dimensions}A and Figure~\ref{fig:dimensions}D).   
% 1. 水资源丰富，耗水量少且灵活用水
In the P1, natural surface water resources were rather abundant with lesser water consumptions (supplementary Fig. S1 and S3) and most of which were flexible water utilization (supplementary Fig. S4). During the P1 and even P2, however, water consumption increases rapidly and natural surface water resources decreases at the same time, making water increasingly scarce. Opposite effect to that, numerous reservoirs built reduced the variability of water resources by boosting storage capacities, but most of which built in P1 (supplementary Fig. S5 and supplementary Table S1). As a result, water utilization stress decreases during P1, but begins to rise rapidly in P2.

% 另一方面，P2-P3，持续变化的水资源利用倾向与格局
On the other hand, as the most positive contributors to the IWRU index in P2 and P3 separately, tendentiousness and patterns of water utilization were keeping to enlarge their impacts (Figure~\ref{fig:dimensions}B and Figure~\ref{fig:dimensions}C). 
% 首先是用水比例的变化
Representing tendentiousness of water utilization, increasing non-provisioning share of water utilization were mainly contributed by larger industrial water consumptions and minor total water uses, while the influence of both are weakening (Figure~\ref{fig:dimensions}E).
% 再讨论用水格局的变化
However, patterns of water utilization, whose contributions to the IWRU are increasing, were mainly benefited from decreasing disparities in the amount of water resources used, both intersectoral and regional (Figure~\ref{fig:dimensions}F).

@@

Some main drivers caused the above changes of water utilization regime.
% 经济总量提升导致资源耗竭的加速。
(1) The expansion of irrigated area and the economic growth of industry and services are keys to the changes in the tendentiousness of water utilization between P1 and P2 (Figure\ref{fig:Causes} A). During the P1, irrigated agricultural area in the Yellow River basin expanded rapidly at a rate of $0.25*10^6/year$, and irrigation water was the dominant utilization way ($81.56\%$ of the total water use in 1965, and $83.17\%$ in 1978). Entering P2, however, while the expansion of irrigated area stalled, industry and services gradually took off and took up more water resources (Figure\ref{fig:Causes} B), leading to $8\%$ reduction of proportion of irrigation water.

@@

% 水资源利用制度包括压力、倾向性、格局三个重要因素，三个指标在过渡框架中随着不断变化。
Throughout the above transition phases towards dualistic water cycle, three dimensions of water utilization regime are various in discipline of evolution, corresponding to typical changes within the Yellow River Basins as an example (Figure~\ref{fig:dimensions}). 
% 水资源压力
(1) Firstly, although stresses on water resources increases by economic expansion boosting water demands, socio-economic progress responding to resource scarcity by better management and efficiency. Water resources were becoming more scarce in the Yellow River Basin from P1 to P3, but the expansion of farmland, the construction of reservoirs, and the increase in water use efficiency became responses to different water utilization stresses in aspects (Figure~\ref{fig:Causes}). Since the scarcity of water resources is directly perceptible and sensitive for utilization, its stresses on societies is one of the most striking drivers to regime shifts within human-water systems \cite{qinFlexibilityIntensityGlobal2019}.
% 水资源倾向性
(2) Secondly, the non-provisioning part of water demands growths with secondary and tertiary industries developing, leading tendentiousness of water utilization continually tilted to the socio-economic part. As original region of Ancient Chinese Civilization, the Yellow River Basin used to be dominated by agricultural but in its way to an energy industry zone now \cite{WillEnergyBases}. As a result, saving water consumption in agriculture and making concessions for industry and energy is widely recognized as solutions for the competing \cite{xiangWillEnergyIndustry2016,bebbWaterRightsTransfers2011}. Anyhow, this changes of tendentiousness reflect a truth that growing socio-economic parts are throwing feedbacks to scarcity of water resources and contributing to regime shifts.
% 水资源格局
(3) Thirdly, with closer socio-economic ties and stiffer competition between regions and sectors, the geographic scope of water resource supply and demand allocation is expanding, leading to changing patterns of water utilization. In the Yellow River Basin, the gap in water consumptions between regions and sectors are narrowing, as the result of a carefully designed allocation \cite{wangThirtyYearsYellow2018}. However, the allocation of water utilization is determined on the basis of regional and sectoral economic contexts and development trajectories\cite{wangThirtyYearsYellow2018}. The changes in water utilization patterns along with regime shifts, therefore, are the outcomes of feedback loops within complex human-water systems.

@@

% 展望：未来还可能陷入结构陷阱，（表现形式：用水效率悖论的同时，倾向性停滞，灵活性进一步降低，）
However, in the regional cycle regime of water utilization, the Yellow River Basins still faces structural traps and require further transformation according to changes within the human-water system. 
Firstly, in line with paradox of irrigation efficiency, significant improvement in agricultural irrigation efficiency (i.e. decline in water use intensity) has been accompanied by a resurgence in irrigated area, resulting in an unabated and weak upward trend in water stress \cite{graftonParadoxIrrigationEfficiency2018}.
% 水资源压力仍在增大，而节水能力已经进入瓶颈，倾向性贡献减少）
Secondly, the changing tendentiousness between non-provisioning (i.e. industry and urban services) and provisioning (i.e. domestic and irrigation water use) is stagnating (see Figure~\ref{fig:dimensions}E) because of rigidify in the industrial structure. 
At the same time, the flexibility of water use is declining since domestic water use and thermal water use growth rapidly (supplementary Fig. S4). 
% 下一步：可以通过统一调控，调整格局来调整结构，突破结构陷阱。
% 总结：整个流域需进一步转型与适应，以实现高质量发展。
Typically, these may lead to a reduction in resilience of basins and leave highly coupled human-water systems facing greater vulnerability to collapse --as a structural trap \cite{cummingResilienceBigRiver2011}. Therefore, based on the identification of current phases and development traps by the transition framework, further transformative governance is still needed to achieve a high-quality sustainable development of the basin \cite{scoonesTransformationsSustainabilityCombining2020a}.

@@

	\subsection*{Integrated Water Resources Utilization (IWRU) Index}
	The Integrated Water Resources Utilization (IWRU) Index consists of three dimensions (stresses, tendentiousnesses and patterns, denoted by sub-indicators correspondingly). Assuming they have equal weights, we log-transformed and standardized the three sub-indicators for elimination of differences between indicators. It means for each indicator $I_{sub}$, we performed:
	$$ I' = log(I_{sub}) $$
	$$ I = (I' - I'_{min}) / (I'_{max}-I'_{min})$$
	where $I$ is standardized series for $I_{sub}$.
	Then, since we assumed different relationships between development and sub-indicators, we added them together in equal weights:
	$$ IWRU = \sum_{i}^3 I'_i $$
	where $i$ is stress, tendentiousness or pattern, and $I'_i$ is standardized sub-indicator of them. A brief description of the sub-indicators used to measure each dimension follows (see SI Methods. S3 for more details).

@@

	\subsubsection*{Sub-indicator of Tendentiousness}
	To tendentiousness, we use non-provisioning shares of water use as an indicator. While provisional water use ($WU_{pro}$) includes domestic, irrigated and livestock water uses, the non-provisioning water use ($WU_{non-pro}$) includes industrial and urban services water uses. Then, we can calculate the non-provisioning shares by:
	$$ NPS_{ij} = \frac{WU_{indirect, i, j}}{WU_{direct, i, j} + WU_{undirect, i, j}} $$

@@

	% 计算贡献度
	\subsection*{Contribution decomposition}
		We have decomposed the amount of variation in each index at different stages in order to observe the contribution of each influencing factor to them. Use Integrated Water Resources Utilization (IWRU) Index as an example, which influenced by three dimensions: stress ($S$), tendentiousness ($T$) and pattern ($P$) (indicated by their own index respectively, see Water utilization regime index and supplementary Methods. S3):
		$$ IWRU = T * P * S ^ {-1} $$
		Take the logarithm of both sides then, we get:
		$$ ln(IWRU) = ln(S) + ln(T) - ln(P) $$
		Since the changes of IWRU $\Delta IWRU$ can be expressed as $\Delta IWRU = ln(IWRU_2) - ln(IWRU_1)$, where $IWRU_2$ and $IWRU_1$ are ending and beginning values of IWRU in a certain time period, combining the above equations we can get:
		$$ \Delta IWRU = ln(\frac{S_1}{S_2}) + ln(\frac{T_2}{T_1}) + ln(\frac{P_2}{P_1}) = -\Delta S + \Delta T + \Delta P $$
		Then, we can calculate contributions $C_F$ of a certain factor $F$ in a certain period by:
		$$ C_{F} = \frac{|\Delta C_{F}|}{\Delta IWRU}$$
}

>>>>

% 文章标题：“流域尺度的水资源利用体系：过渡框架和发展困局”
\title{Universal regime transition for water resources utilization in developing river basins}

@@

%tag 摘要
\begin{abstract}
	From agrarian to industrial societies, humans have harnessed ecosystems over rivers basins around the world, while using water resources as blood that sustains social developments. 
	Although relations between societies and water resources keep changing throughout, there is still lacking effective methods to detect related regimes of water utilization, with much fewer attempts to develop theoretical framework to explain their transitions as well. 
	Here, by integrating three equally important but differentiated dimensions of water utilization (stress, priority and lopsidedness), we develop an Integrated Water Resources Utilization (IWRU) Index at a basin scale to indicate regime shifts. 
	By applying this index to the Yellow River Basin, China, results suggest three water utilization regimes in developing over half a century, whose shifts led by various but pervasive causes.
	Based on that, we summarized a general transition framework which gives a sketch of relationships between human societies and their water utilization, as a useful guideline for big river basins to develop in a coordinated way.
\end{abstract}

@@

% Tag 引言第三段
% 前人已经从不同维度刻画了人水关系.
Development of societies by using water resources has been going on for at least thousands of years. Although its regime shifts and transition phases are not fully understood yet, water utilization has been depicted and studied from different dimensions.
% 首先，因为水资源的稀缺性和全球用水量的增加，受到最广泛关注的是人类社会面临的水资源压力。
Firstly, since water resources are scarce, the most widespread concern is the rising stresses on human societies to use water resources.
Even though the stocks of water in artificial reservoirs are helpful to water resources availability, greater water utilization stresses had become a major constraint to development, because of significant increment in water withdrawals and larger shares of inflexible water utilization during the last century.
\cite{postelHumanAppropriationRenewable1996, greveGlobalAssessmentWater2018a, qinFlexibilityIntensityGlobal2019}
% 其次，随着工业迅猛发展和生态建设的需要，社会对水资源的利用的倾向性也发生了转变。
Secondly, as the need of industrial and ecological developments, priority of water utilization changed with. 
Despite a major water utilization of agricultural irrigation dominating most river basins, there are noticeable growths and preferential priority in the economy profits and water consumption regarding industry or services, leading potential conflicts between different sectors.
\cite{liuWaterScarcityAssessments2017, florkeWaterCompetitionCities2018}
% 最后，由于水的可用性本质上是区域问题，水资源利用的格局也很重要
Thirdly, since water distribution and utilization are inherently basinal concerns, patterns of also play an important role.
% 全球总的取水量很少，但缺水地区很多
Although only 10\% of available water is withdrawn on global average, about 30\% of population live in highly water-stressed areas,
where dominated sectors of water utilization are various as well. 
\cite{wadaWedgeApproachWater2014, okiGlobalHydrologicalCycles2006}
% 此外，人类活动还在改变这一格局
In addition, human activities are still changing this pattern, since positive impacts caused by human interventions mostly occur in upper regions whereas aggravated water resources downstream, in many basins around the world.
\cite{veldkampWaterScarcityHotspots2017}
% 将三种视角结合，就是“水资源利用体系”。
Although existing researches have evaluated the aspects of water resource utilization from these different dimensions, we still cannot obtain a coherent understanding of regime shifts regard to social development and water utilization, without integrating them.

@@

\section*{Results}
\subsection*{Detection of Water utilization regimes}
% 这一节主要展示IWRU的变化趋势和WUR的划分
With two significant points detected, the trend of IWRU index are split into three periods, whose slopes are various and mainly contributed by different dimensions (stress, priority or lopsidedness of water utilization, see Methods) (Figure~\ref{fig:IWRU}).
\begin{figure}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/index.pdf}
	\caption{Changes of the IWRU index. 
	\textbf{A,} with two change points in 1978 and 1994, three periods were detected in trend of the IWRU.
	\textbf{B,} changes of IWRU in three periods have various slopes, while the second period have a negative growths rate.
	\textbf{C,} changes of the IWRU within three certain periods, which have different main contributors.
	}
	\label{fig:IWRU}
\end{figure}
% 接下来分句介绍每个阶段的特征
% 第一阶段
In the first period (P1, 1965-1978), the IWRU index had a rapidly increasing and the lightening of water stresses made the most striking contribution (+0.722), while priority and lopsidedness of the water utilization had slight negative contribution (-0.048 and -0.09 respectively).
% 第二阶段
In the second period (P2, 1979-1994), the IWRU index experienced a slight drop, despite positive contributions of priority and lopsidedness of water utilization (+0.352 and +0.279 respectively), because of increasing stresses on water resource playing a larger negative role (-0.636). 
% 第三阶段
However, as the further increasing of positive contributions of water utilization priority (+0.485) and lopsidedness (+0.515), and decelerations of water stresses (-0.344, 46\% less than P2) in the third period (P3, 1995-2013), a positive growth of the IWRU returned.
% 总而言之，每个阶段都由水资源利用的不同维度提供最大的正向作用
As a result, each period has a different most striking positive contributor to IWRU: P1 is stress; P2 is priority; and P3 is lopsidedness.

% 用水三个维度的组合呈现出阶段特征明显
Combining these three dimensions' net contribution to IWRU further, ratios of the contributions of the three dimensions clustered clearly by different time periods, indicating three regimes (Figure~\ref{fig:phases}).
\begin{figure}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/phases.pdf}
	\caption{Combination of three dimensions (S: stresses; T: priority; P: lopsidedness) in different periods. 
	Size of the points denoting average of absolute value of each indicator: the mean of the P1 phase is 0.10, while 0.14, 0.19 in P2 and P3.
	The red indicator line in this ternary plot denotes 1:1 contributions between priority (T) and patterns (P).
	Two key change points (1978, 1994), along with the beginning (1965) and the ending (2013) of research period, are labelled.}
	\label{fig:phases}
\end{figure}
% 第一阶段到第二阶段
At the very beginning (1965) and throughout the whole P1, water utilization regime dominated by high stress. After then, it experienced a shift to low stress since 1978, with a change in the proportion of contributions between priority and lopsidedness, too.
% 第三阶段集中
Finally, the three dimensions' contribution were much similar in P3, making the points highly concentrated at the centre of the ternary diagram for that period.

%tag 结果2
\subsection*{Differences between water utilization regimes}
% 在不同的维度下各阶段进行对比，可以发现不同的水资源利用体系间存在明显差异。
The differences between the water utilization regimes are reflected in changes of all three dimensions.
\begin{figure*}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/dimensions.pdf}
	\caption{
		Changes in different dimensions of water resources utilization regimes and their main contributors.
		\textbf{A,} changes of water utilization stress, indicated by unstandardized scarcity-flexibility-variability water stresses index (SFV-index).
		\textbf{B,} changes of water utilization priority, indicated by non-provisioning water shares.
		\textbf{C,} changes of water utilization lopsidedness, indicated by unstandardized distribution information entropy index (see supplementary: Methods. S3 for details).
		\textbf{D,} Main impact factors to water utilization stresses in each period or region, and their contributions to changes of unstandardized SFV-index.
		\textbf{E,} Main impact water uses to water utilization tendentiousnesses, and their contributions to changes of non-provisioning water shares.
		\textbf{F,} Main impact factors to water utilization patterns, and their contributions to changes of related unstandardized distribution information entropy index (see supplementary: Methods. S3 for details).
		}
	\label{fig:dimensions}
\end{figure*}
% 从阶段一到阶段二，水资源压力的变化最为明显
Moving from the regime in P1 to P2, the most striking change is the reversal of the trend in water utilization stress, which is determined by a combination of scarcity, flexibility and variability (Figure~\ref{fig:dimensions}A and Figure~\ref{fig:dimensions}D).   
% 1. 水资源丰富，耗水量少且灵活用水
In the P1, natural surface water resources were rather abundant with lesser water consumptions (supplementary Fig. S1 and S3) and most of which were flexible water utilization (supplementary Fig. S4). During the P1 and even P2, however, water consumption increases rapidly and natural surface water resources decreases at the same time, making water increasingly scarce. Opposite effect to that, numerous reservoirs built reduced the variability of water resources by boosting storage capacities, but most of which built in P1 (supplementary Fig. S5 and supplementary Table S1). As a result, water utilization stress decreases during P1, but begins to rise rapidly in P2.

% 另一方面，P2-P3，持续变化的水资源利用倾向与格局
On the other hand, as the most positive contributors to the IWRU index in P2 and P3 separately, priority and patterns of water utilization were keeping to enlarge their impacts (Figure~\ref{fig:dimensions}B and Figure~\ref{fig:dimensions}C). 
% 首先是用水比例的变化
Representing priority of water utilization, increasing non-provisioning share of water utilization were mainly contributed by larger industrial water consumptions and minor total water uses, while the influence of both are weakening (Figure~\ref{fig:dimensions}E).
% 再讨论用水格局的变化
However, patterns of water utilization, whose contributions to the IWRU are increasing, were mainly benefited from decreasing disparities in the amount of water resources used, both intersectoral and regional (Figure~\ref{fig:dimensions}F).

@@

Some main drivers caused the above changes of water utilization regime.
% 经济总量提升导致资源耗竭的加速。
(1) The expansion of irrigated area and the economic growth of industry and services are keys to the changes in the priority of water utilization between P1 and P2 (Figure\ref{fig:Causes} A). During the P1, irrigated agricultural area in the Yellow River basin expanded rapidly at a rate of $0.25*10^6/year$, and irrigation water was the dominant utilization way ($81.56\%$ of the total water use in 1965, and $83.17\%$ in 1978). Entering P2, however, while the expansion of irrigated area stalled, industry and services gradually took off and took up more water resources (Figure\ref{fig:Causes} B), leading to $8\%$ reduction of proportion of irrigation water.

@@

% 水资源利用制度包括压力、倾向性、格局三个重要因素，三个指标在过渡框架中随着不断变化。
Throughout the above transition phases towards dualistic water cycle, three dimensions of water utilization regime are various in discipline of evolution, corresponding to typical changes within the Yellow River Basins as an example (Figure~\ref{fig:dimensions}). 
% 水资源压力
(1) Firstly, although stresses on water resources increases by economic expansion boosting water demands, socio-economic progress responding to resource scarcity by better management and efficiency. Water resources were becoming more scarce in the Yellow River Basin from P1 to P3, but the expansion of farmland, the construction of reservoirs, and the increase in water use efficiency became responses to different water utilization stresses in aspects (Figure~\ref{fig:Causes}). Since the scarcity of water resources is directly perceptible and sensitive for utilization, its stresses on societies is one of the most striking drivers to regime shifts within human-water systems \cite{qinFlexibilityIntensityGlobal2019}.
% 水资源倾向性
(2) Secondly, the non-provisioning part of water demands growths with secondary and tertiary industries developing, leading priority of water utilization continually tilted to the socio-economic part. As original region of Ancient Chinese Civilization, the Yellow River Basin used to be dominated by agricultural but in its way to an energy industry zone now \cite{WillEnergyBases}. As a result, saving water consumption in agriculture and making concessions for industry and energy is widely recognized as solutions for the competing \cite{xiangWillEnergyIndustry2016,bebbWaterRightsTransfers2011}. Anyhow, this changes of priority reflect a truth that growing socio-economic parts are throwing feedbacks to scarcity of water resources and contributing to regime shifts.
% 水资源格局
(3) Thirdly, with closer socio-economic ties and stiffer competition between regions and sectors, the geographic scope of water resource supply and demand allocation is expanding, leading to changing patterns of water utilization. In the Yellow River Basin, the gap in water consumptions between regions and sectors are narrowing, as the result of a carefully designed allocation \cite{wangThirtyYearsYellow2018}. However, the allocation of water utilization is determined on the basis of regional and sectoral economic contexts and development trajectories\cite{wangThirtyYearsYellow2018}. The changes in water utilization patterns along with regime shifts, therefore, are the outcomes of feedback loops within complex human-water systems.

@@

% 展望：未来还可能陷入结构陷阱，（表现形式：用水效率悖论的同时，倾向性停滞，灵活性进一步降低，）
However, in the regional cycle regime of water utilization, the Yellow River Basins still faces structural traps and require further transformation according to changes within the human-water system. 
Firstly, in line with paradox of irrigation efficiency, significant improvement in agricultural irrigation efficiency (i.e. decline in water use intensity) has been accompanied by a resurgence in irrigated area, resulting in an unabated and weak upward trend in water stress \cite{graftonParadoxIrrigationEfficiency2018}.
% 水资源压力仍在增大，而节水能力已经进入瓶颈，倾向性贡献减少）
Secondly, the changing priority between non-provisioning (i.e. industry and urban services) and provisioning (i.e. domestic and irrigation water use) is stagnating (see Figure~\ref{fig:dimensions}E) because of rigidify in the industrial structure. 
At the same time, the flexibility of water use is declining since domestic water use and thermal water use growth rapidly (supplementary Fig. S4). 
% 下一步：可以通过统一调控，调整格局来调整结构，突破结构陷阱。
% 总结：整个流域需进一步转型与适应，以实现高质量发展。
Typically, these may lead to a reduction in resilience of basins and leave highly coupled human-water systems facing greater vulnerability to collapse --as a structural trap \cite{cummingResilienceBigRiver2011}. Therefore, based on the identification of current phases and development traps by the transition framework, further transformative governance is still needed to achieve a high-quality sustainable development of the basin \cite{scoonesTransformationsSustainabilityCombining2020a}.

@@

	\subsection*{Integrated Water Resources Utilization (IWRU) Index}
	The Integrated Water Resources Utilization (IWRU) Index consists of three dimensions (stresses, tendentiousnesses and patterns, denoted by sub-indicators correspondingly). Assuming they have equal weights, we log-transformed and standardized the three sub-indicators for elimination of differences between indicators. It means for each indicator $I_{sub}$, we performed:
	$$ I' = log(I_{sub}) $$
	$$ I = (I' - I'_{min}) / (I'_{max}-I'_{min})$$
	where $I$ is standardized series for $I_{sub}$.
	Then, since we assumed different relationships between development and sub-indicators, we added them together in equal weights:
	$$ IWRU = \sum_{i}^3 I'_i $$
	where $i$ is stress, priority or lopsidedness, and $I'_i$ is standardized sub-indicator of them. A brief description of the sub-indicators used to measure each dimension follows (see SI Methods. S3 for more details).

@@

	\subsubsection*{Sub-indicator of priority}
	To priority, we use non-provisioning shares of water use as an indicator. While provisional water use ($WU_{pro}$) includes domestic, irrigated and livestock water uses, the non-provisioning water use ($WU_{non-pro}$) includes industrial and urban services water uses. Then, we can calculate the non-provisioning shares by:
	$$ NPS_{ij} = \frac{WU_{indirect, i, j}}{WU_{direct, i, j} + WU_{undirect, i, j}} $$

@@

	% 计算贡献度
	\subsection*{Contribution decomposition}
		We have decomposed the amount of variation in each index at different stages in order to observe the contribution of each influencing factor to them. Use Integrated Water Resources Utilization (IWRU) Index as an example, which influenced by three dimensions: stress ($S$), priority ($T$) and lopsidedness ($P$) (indicated by their own index respectively, see Water utilization regime index and supplementary Methods. S3):
		$$ IWRU = T * P * S ^ {-1} $$
		Take the logarithm of both sides then, we get:
		$$ ln(IWRU) = ln(S) + ln(T) - ln(P) $$
		Since the changes of IWRU $\Delta IWRU$ can be expressed as $\Delta IWRU = ln(IWRU_2) - ln(IWRU_1)$, where $IWRU_2$ and $IWRU_1$ are ending and beginning values of IWRU in a certain time period, combining the above equations we can get:
		$$ \Delta IWRU = ln(\frac{S_1}{S_2}) + ln(\frac{T_2}{T_1}) + ln(\frac{P_2}{P_1}) = -\Delta S + \Delta T + \Delta P $$
		Then, we can calculate contributions $C_F$ of a certain factor $F$ in a certain period by:
		$$ C_{F} = \frac{|\Delta C_{F}|}{\Delta IWRU}$$
}
